Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-09-07 paper/main.tex
[ paper ] new commands for programs and langs.
minor changes.
[ paper ] section 2 improvements.
[ paper ] major changes, abstract, conclusions was added.

\usepackage{xargs}
\usepackage[pdftex,dvipsnames]{xcolor}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]
  {\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]
  {\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]
  {\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25
        ,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]
  {\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]
  {\todo[disable,#1]{#2}}

@@

% --------------------------------------------------------------------------
% New commands
% --------------------------------------------------------------------------

\newcommand{\name}{\textsc}
\newcommand{\Agda}{\name{Agda}}

@@

\title{Towards Proof-Reconstruction of Problems
in Classical Propositional Logic in Agda}

@@

\institute{Universidad EAFIT, \emph{School of Sciences and Engineering},\\
Medell\'in, Colombia\\
\email{jprieto9,asr@eafit.edu.co}}

@@

\subfile{TODO.tex}

@@

\subfile{sections/abstract.tex}
\subfile{sections/introduction.tex}
\subfile{sections/metis-language-and-proofs.tex}
\subfile{sections/proof-reconstruction.tex}
\subfile{sections/related-work.tex}
\subfile{sections/conclusions.tex}
\subfile{sections/acknowledgements.tex}

>>>>

\usepackage{xargs}
\usepackage[dvipsnames]{xcolor}
% \usepackage[pdftex,dvipsnames]{xcolor}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]
  {\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]
  {\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]
  {\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25
        ,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]
  {\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]
  {\todo[disable,#1]{#2}}

@@

% --------------------------------------------------------------------------
% New commands
% --------------------------------------------------------------------------
\usepackage{xspace}

\newcommand{\abbre}[1]{\textsc{#1}\xspace}
\newcommand{\ATP}{\abbre{ATP}}
\newcommand{\ATPs}{\abbre{ATPs}}
\newcommand{\SAT}{\abbre{SAT}}
\newcommand{\SMT}{\abbre{SMT}}
\newcommand{\ITP}{\abbre{ITP}}
\newcommand{\CPL}{\abbre{CPL}}
\newcommand{\CNF}{\abbre{CNF}}

\newcommand{\name}[1]{\texttt{#1}\xspace}
\newcommand{\canonicalize}{\name{canonicalize}}
\newcommand{\clausify}{\name{clausify}}
\newcommand{\conjunct}{\name{conjunct}}
\newcommand{\negate}{\name{negate}}
\newcommand{\resolve}{\name{resolve}}
\newcommand{\simplify}{\name{simplify}}
\newcommand{\strip}{\name{strip}}
\newcommand{\tptpaxiom}{\name{axiom}}
\newcommand{\tptpconjecture}{\name{conjecture}}

\newcommand{\prg}[1]{\texttt{#1}\xspace}
\newcommand{\Agda}{\prg{Agda}}
\newcommand{\Metis}{\prg{Metis}}
\newcommand{\IsabelleHOL}{\prg{IsabelleHOL}}

\newcommand{\len}[1]{\texttt{#1}\xspace}
\newcommand{\Haskell}{\len{Haskell}}
\newcommand{\TPTP}{\len{TPTP}}
\newcommand{\TSTP}{\len{TSTP}}

@@

\title{Towards Proof-Reconstruction of Problems
in Classical Propositional Logic}

@@

\institute{Universidad EAFIT, Logic and Computation Research Group,\\
Department of Mathematics, Medell\'in, Colombia\\
\email{\{jprieto9,asr\}@eafit.edu.co}}

@@

@@

\subfile{sections/abstract.tex}
\subfile{sections/introduction.tex}
\subfile{sections/metis-language-and-proofs.tex}
% \subfile{sections/proof-reconstruction.tex}
\subfile{sections/related-work.tex}
\subfile{sections/conclusions.tex}
\subfile{sections/acknowledgements.tex}

### 2017-09-07 paper/sections/abstract.tex
[ paper ] new commands for programs and langs.
[ paper ] major changes, abstract, conclusions was added.

% ============================================================================

\begin{abstract}
...
\keywords{Proof reconstruction, Agda, Automatic theorem prover,
  Classical Propositional Logic, Metis }
\end{abstract}

>>>>

% ===================================================================

\begin{abstract}
In this paper, we describe our approach to reconstruct and verify in
\Agda, a  proof assistant,  solutions obtained by \Metis, an
automatic theorem prover, to problems in classical propositional
logic. The translation method consists of justifying step-by-step
each deduction found in the prover's solution. To carry out this
plan, we emulate the inference rules of the prover, providing a
version of them in the proof assistant. Finally, we parse the output
of the prover,  constructing the respective tree associated with the
proof, ending up with an \Agda code for type-checking.

\keywords{proof reconstruction, \Agda, automatic theorem prover, classical propositional logic, \Metis.}
\end{abstract}

### 2017-09-07 paper/sections/acknowledgements.tex
[ paper ] new commands for programs and langs.
[ paper ] major changes, abstract, conclusions was added.

% ============================================================================

@@

We thank Universidad EAFIT of Medell\'in, Colombia for funding support.
This project was a product of the Logic and Computation Research
Group at the Universidad EAFIT.
%This is part of first author's Master thesis in Applied Mathematics, written under the
% supervision of Andr\'es Sicard-Ram\'irez for the Logic and Computation Research
% Group at the Universidad EAFIT.

We thank Joe Leslie-Hurd for all support with \verb!Metis! prover.
We also acknowledge the work carried out by Alejandro G\'omez-Londo\~no~
\cite{Gomez-Londono2015} used as the basis code for our TSTP parsing
module.
And last but not least, we thankfully acknowledge Andreas Abel and Chalmers
University of Goteborg, Sweden for inviting us to be part of the Agda
Implementors’ Meeting XXV where we presented part of this paper.

>>>>

% ===================================================================

@@

We thank Universidad EAFIT of Medell\'in, Colombia for funding
support. This project was a product of the Logic and Computation
Research Group at the Universidad EAFIT.

We thank Joe Leslie-Hurd for his support with \Metis prover. We also
acknowledge the work carried out by Alejandro G\'omez-Londo\~no~
\cite{Gomez-Londono2015} used as the basis code for our \TSTP
parsing module.
\improvement{We should consider thank Kanso and Anton, to provide us
their thesis, discussion, and source code to compare with our work?}

Last but not least, we thankfully acknowledge
Andreas Abel and Chalmers University of Goteborg, Sweden for
inviting us to be part of the \Agda Implementors’ Meeting XXV where
we presented part of this paper.

### 2017-09-07 paper/sections/conclusions.tex
[ paper ] major changes, abstract, conclusions was added.

Amazing conclusions.
...
\cite{}

\subsubsection*{Future Work.}
\label{ssec:future-work}

>>>>

Besides our scope in this document, propositional logic without
equality, is straightforward compared with other logics like
first-order or high-order logic; the proof reconstruction task is
far from trivial. Depending on the systems involved in the
translation, some difficulties can arise. We listed some of them in
section \ref{sec:metis-language-and-proofs} and section
\ref{sec:proof-reconstruction}.

For instance, just showing equality between Normal Forms for the
same formula in the output of the \canonicalize inference and a
common implementation for this kind of conversion; we observed that
\Metis performs reordering for each expression mainly in the
presence of negative literals. Another example was the \Metis'
convention for conjunctions and disjunction as right associative and
binary operators in the \TSTP format. We had to update the \TSTP
parser module introduced in \cite{Gomez-Londono2015} to include some
feature like this.

However, we believe one of our biggest obstacles was dealing with
the Equality between formulas since, in our approach, we only work
with the syntactic treatment of the logic without considering the
semantics. One of the main reasons for such a restriction resides in
our plans to support further first-order logic where a semantic
treatment is not viable.\unsure{Please confirm this assumption.}

Nevertheless, the proof reconstruction is certainly possible, and
the verification of solutions of \CPL problems can be carried out
successful in \Agda. Our tool is capable of translating \Metis
derivations in \TSTP format to deliver \Agda code of the proof, and
this proof can be type-checked by \Agda.

We believe the source code of our tool can be extended to support
other \ATPs for \CPL like \name{E}. For example, we incorporate the
inferences rules of \name{E} or \name{Z3} from the description in
Kanso's Thesis~\cite{Kanso2012}.

Trying to justify output from a theorem prover has a real impact,
increasing the trustworthy of the prover. For instance, we found a
soundness bug, a bug in the printing out of the \TSTP derivations,
and a bug in the parsing for \TPTP problem in \Metis.
Fortunately, all these bugs were fixed quickly by Hurd\footnote{
Issues No. 4, No. 2, and the commit \name{8a3f11e} from the
\Metis' official repository \url{https://github.com/gilith/metis}},
and a new version of \Metis was released
(version 2.3, release 20170810).

% \subsubsection*{Future Work.}
% \label{ssec:future-work}

### 2017-09-07 paper/sections/introduction.tex
[ paper ] major changes, abstract, conclusions was added.

Proof reconstruction is a hard labor since it depends on the integration
of two complex systems. On the one hand, we have the automatic theorem provers
(henceforth \ATP) and their specification logic. These tools are usually
classified in at least one of the following categories. A \SAT solver
(e.g., \name{zChaff}~\cite{Moskewicz2001} and \name{MiniSat}~\cite{Een2004})
to prove unsatisfiability of \CNF formulas, a \abbre{QBF} solver
(e.g., \name{GhostQ}~\cite{Klieber2014} and \name{DepQBF}~\cite{Lonsing2017})
to prove satisfiability and invalidity of quantified Boolean formulas, a \SMT
solver (e.g \name{CVC4}~\cite{Barrett2011}, \name{veriT}~\cite{bouton2009},
and \name{Z3}~\cite{DeMoura2008}) to prove unsatisfiability of formulas
from first-order logic with theories, and a prover for validity of
formulas from first-order logic with equality
(e.g., \name{E}~\cite{Schulz:AICOM-2002}, \name{leanCoP}~\cite{Otten2008},
\Metis~\cite{hurd2003first}, \name{SPASS} ~\cite{Weidenbach2009} and
\name{Vampire}~\cite{Riazanov1999}), high-order logic (e.g., \name{Leo-II}
\cite{Benzmuller2008} and \name{Satallax}~\cite{Brown2012}) or intuitionistic
logic (e.g., \name{ileanCoP}~\cite{Otten2008}, \name{JProver}
\cite{Schmitt2001}, and \name{Gandalf}~\cite{Tammet1997}), among others.
On the other hand, we have the proof checkers, interactive theorem provers
(henceforth \ITP) or proof assistants (e.g., \Agda~\cite{agdateam},
\name{Coq}~\cite{coqteam}, \name{Isabelle}~\cite{paulson1994isabelle}, and
\name{HOL4}~\cite{norrish2007hol}).
The \ITP tools provide us the logic framework to check and validate the
reply of the \ATPs since they allow us to define the formal language for
the problems like operators, logic variables, axioms, and theorems.

A proof reconstruction tool provides such an integration translating 
the reply delivered by the prover into the formalism of
the proof assistant. Because the formalism of the source (the evidence
generated by the \ATP) is not necessarily the same logic in the target, the
reconstruction turns out in a ``reverse engineering'' task.
Then, reconstructing a proof involves an in-depth understanding of the
algorithms in the \ATP and the specification logic in the \ITP.
What we need from the \ATP tools is a proof-object in a consistent format
to work with, that is, a full script describing step-by-step with
exhaustive details and without ambiguities, the table of the derivation
to get the actual proof. For problems in classical propositional logic
(henceforth \CPL), from a list of at least forty\footnote{\ATPs available
from the web service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just
a few provers were able to deliver proofs (e.g., \name{CVC4}~
\cite{Barrett2011}, \name{SPASS}, and \name{Waldmeister}~\cite{hillenbrand1997})
and a little bit less replied with a proof in a file format like
\TSTP~\cite{sutcliffe2004tstp} (e.g., \name{E}, \Metis,
\name{Vampire}, and \name{Z3}), \len{LFSC}~\cite{Stump2008}, or the
\len{SMT-LIB}~\cite{Bohme2011} format.

>>>>

Proof reconstruction is a hard labor since it depends on the
integration of two complex systems. On the one hand, we have the
automatic theorem provers (henceforth \ATP) and their specification
logic. These tools are usually classified in at least one of the
following categories. A \SAT solver (e.g., \name{zChaff}
\cite{Moskewicz2001} and \name{MiniSat}~\cite{Een2004}) to prove
unsatisfiability of \CNF formulas, a \abbre{QBF} solver (e.g.,
\name{GhostQ}~\cite{Klieber2014} and \name{DepQBF}
\cite{Lonsing2017}) to prove satisfiability and invalidity of
quantified Boolean formulas, a \SMT solver (e.g \name{CVC4}
\cite{Barrett2011}, \name{veriT}~\cite{bouton2009}, and \name{Z3}
\cite{DeMoura2008}) to prove unsatisfiability of formulas from
first-order logic with theories, and a prover for validity of
formulas from first-order logic with equality (e.g., \name{E}
\cite{Schulz:AICOM-2002}, \name{leanCoP}~\cite{Otten2008}, \Metis
\cite{hurd2003first}, \name{SPASS} ~\cite{Weidenbach2009} and
\name{Vampire}~\cite{Riazanov1999}), high-order logic (e.g.,
\name{Leo-II} \cite{Benzmuller2008} and \name{Satallax}
\cite{Brown2012}) or intuitionistic logic (e.g., \name{ileanCoP}
\cite{Otten2008}, \name{JProver} \cite{Schmitt2001}, and
\name{Gandalf}~\cite{Tammet1997}), among others. On the other hand,
we have the proof checkers, interactive theorem provers (henceforth
\ITP) or proof assistants (e.g., \Agda~\cite{agdateam}, \name{Coq}
\cite{coqteam}, \name{Isabelle}~\cite{paulson1994isabelle}, and
\name{HOL4}~\cite{norrish2007hol}). The \ITP tools provide us the
logic framework to check and validate the reply of the \ATPs since
they allow us to define the formal language for the problems like
operators, logic variables, axioms, and theorems.

A proof reconstruction tool provides such an integration translating
the reply delivered by the prover into the formalism of the proof
assistant. Because the formalism of the source (the evidence
generated by the \ATP) is not necessarily the same logic in the
target, the reconstruction turns out in a ``reverse engineering''
task. Then, reconstructing a proof involves an in-depth
understanding of the algorithms in the \ATP and the specification
logic in the \ITP. What we need from the \ATP tools is a
proof-object in a consistent format to work with, that is, a full
script describing step-by-step with exhaustive details and without
ambiguities, the table of the derivation to get the actual proof.
For problems in classical propositional logic (henceforth \CPL),
from a list of at least forty\footnote{\ATPs available from the web
service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just a few
provers were able to deliver proofs (e.g., \name{CVC4}
\cite{Barrett2011}, \name{SPASS}, and \name{Waldmeister}
\cite{hillenbrand1997}) and a little bit less replied with a proof
in a file format like \TSTP~\cite{sutcliffe2004tstp} (e.g.,
\name{E}, \Metis, \name{Vampire}, and \name{Z3}), \len{LFSC}
\cite{Stump2008}, or the \len{SMT-LIB}~\cite{Bohme2011} format.

### 2017-09-07 paper/sections/metis-language-and-proofs.tex
[ paper ] major changes, abstract, conclusions was added.

% ===========================================================================

@@

\Metis is an automatic theorem prover written in \len{Standard ML} for
first-order logic with equality developed by John Hurd~\cite{hurd2003first}.
This prover is suitable for proof reconstruction since it provides
well-documented proof-objects to justify its deduction steps from the basis
of only six inference rules. Exists tools that perform proof
reconstruction with \Metis, some of these systems are the following.

@@

% ---------------------------------------------------------------------------

@@

The \name{source} field is an inference record with the following pattern,
\begin{verbatim}
inference(rule, useful info, parents).
\end{verbatim}

The \name{rule} stands for the inference name; the other fields are
supporting arguments or useful information to apply the reasoning step, and
the list of parents nodes.
For instance, in Fig.~\ref{fig:metis-proof-tstp}, \strip is the name of one
of the inference with no arguments and only one parent node, the \name{goal}.

@@

A proof-object delivered in a \Metis' proof encodes a natural
deduction proof. The deduction system uses six inference rules
\cite{hurd2003first} listed in Fig.~\ref{fig:metis-inferences}.
With these inference rules as the only valid deduction steps, \Metis attempts to
prove conjectures by refutation (see. e.g., \emph{falsium} in the last conclusion
of the \TSTP derivation in Fig.~\ref{fig:metis-example}).

@@

We noted in \TSTP derivations that \Metis do not take into account some
inference rules for the \Metis' official documentation; we refer to
the \Metis' logic kernel. Besides the figure that shows us the six
\Metis' inference rules, its proof-objects, the \TSTP derivations,
exhibit an entirely different set of rules.

For instance, for the propositional logic fragment without equality,
we know only three rules are valid from the \Metis' Logic Kernel:
\emph{assume}, \emph{axiom}, and \emph{resolve}.
Nevertheless, as we observed in \TSTP derivations, for such a propositional
fragment, \Metis uses in his deductions steps, six rules where only the
resolve inference persists from the original list. In the case of \resolve,
it differs from its original version since, in \TSTP,  this rules comes with
an argument, a literal. The other undocumented rules found in \TSTP
derivations are: \canonicalize, \clausify, \conjunct, \negate, \simplify,
and \strip.

This lack of the \Metis' documentation was an important matter for us in the
proof reconstruction and certainty for other implementors (see a comment
related on the presenting of \name{Sledgehammer}
\info{cite here Paulson and Metis Paramodulation' paper}).
We faced the ambiguity and lack of clearness of some rules like \canonicalize
or \simplify. For example, we believe that with these last inferences, \Metis
summarized many essential steps in the conversion or deduction from one formula
or a list of formulas to another as we show in Example X.
\unsure{Should I give an Example?}
That is inconvenient since, our proof reconstruction approach needs take into
account any assumption or consideration as ordering for literals, application
of simplification theorems, reductions, among other valid transformations.
\unsure{Should I give an Example?}
In this part, our job\improvement{job or purpose?} mainly consists of giving
a response to these rules looking at the \Metis' source code while inspecting
a representative set of solutions of \CPL problems generated by \Metis.
\unsure{Should I mention the Prop-Pack Collection?}
Therefore, this is an attempt to help and fulfill this difficulty, describing
our understanding of each rule, to later on in section
\ref{sec:proof-reconstruction}, formalize them effectively in \Agda.

% ...........................................................................

@@

% ...........................................................................

@@

% ...........................................................................

\textit{Normalization.}
\Metis uses the \canonicalize rule to normalize a
propositional formula that comes from an axiom or a definition.
It often converts each expression in one of its normal form,
the conjunctive normal form, the negative normal form or the disjunctive normal form.
Posterior, this rule simplifies the formula with some definitions, applying
recursively on the structure's formula conjunctions and disjunctions simplifications
to remove tautologies
(see Fig.~\ref{fig:conjunctive-disjunctive-simplification} with a list
of these theorems. We had assumed the commutative property for conjunction and
disjunction connectives).

@@

\[%\scalebox{0.9}{
  \begin{bprooftree}
    \AxiomC{$\varphi \vee \bot$}
    \UnaryInfC{$\varphi$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{$\varphi \vee \top$}
    \UnaryInfC{$\top$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{$\varphi \vee \neg \varphi$}
    \UnaryInfC{$\top$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{$\varphi \vee \varphi$}
    \UnaryInfC{$\varphi$}
  \end{bprooftree}
%}
\]
\caption{Defintions of \canonicalize inference.}
\label{fig:conjunctive-disjunctive-simplification}
\end{figure}

% ...........................................................................

@@

% ...........................................................................

\textit{Simplification.} %abbreviate formulas
The \simplify rule is a $n$-ary inference that performs simplification
of definitions. This rule transverses a list of formulas by
applying different theorems (e.g., \emph{modus pones}, \emph{modus tollens},
or \emph{disjunctive syllogism}) including the list of theorems used by
\canonicalize in Fig.~\ref{fig:conjunctive-disjunctive-simplification}
and the resolution theorem of \resolve rule.
Moreover, many things happen inside this \Metis' procedure, and
it is somewhat complicated to grasp it completely.

>>>>

% ===================================================================

@@

\Metis is an automatic theorem prover written in \len{Standard ML}
for first-order logic with equality developed by John Hurd
\cite{hurd2003first}. This prover is suitable for proof
reconstruction since it provides well-documented proof-objects to
justify its deduction steps from the basis of only six inference
rules. Exists tools that perform proof reconstruction with \Metis,
some of these systems are the following.

@@

% -------------------------------------------------------------------

@@

The \name{source} field is an inference record with the following
pattern,

\begin{verbatim}
inference(rule, useful info, parents).
\end{verbatim}

The \name{rule} stands for the inference name; the other fields are
supporting arguments or useful information to apply the reasoning
step, and the list of parents nodes. For instance, in Fig.
\ref{fig:metis-proof-tstp}, \strip is the name of one of the
inference with no arguments and only one parent node, the
\name{goal}.

@@

A proof-object delivered in a \Metis' proof encodes a natural
deduction proof. The deduction system uses six inference rules
\cite{hurd2003first} listed in Fig.~\ref{fig:metis-inferences}.
With these inference rules as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (see. e.g.,
\emph{falsium} in the last conclusion of the \TSTP derivation in
Fig.\ref{fig:metis-example}).

@@

We noted in \TSTP derivations that \Metis do not take into account
some inference rules for the \Metis' official documentation; we
refer to the \Metis' logic kernel.
Besides the six \Metis' inference rules from the Fig.
\ref{fig:metis-inferences}, its proof-objects, the \TSTP
derivations, exhibit an entirely different set of rules.

For instance, for the propositional logic fragment without equality,
we know only three rules are valid from the \Metis' Logic Kernel:
\emph{assume}, \emph{axiom}, and \emph{resolve}. Nevertheless, as we
observed in \TSTP derivations, for such a propositional fragment,
\Metis uses in his deductions steps, six rules where only the
\emph{resolve} inference persists from the original list. In the
case of \resolve, it differs from its original version since, in
\TSTP, this rules comes with an argument, a literal. The other
undocumented rules found in \TSTP derivations are: \canonicalize,
\clausify, \conjunct, \negate, \simplify, and \strip.

This lack of the \Metis' documentation was an important matter for
us in the proof reconstruction and certainty for other implementors
(see a comment related on the presenting of \name{Sledgehammer}
in \cite{paulson2007source} and \cite{Farber2015}). We faced
the ambiguity and lack of clearness of some rules like \canonicalize
or \simplify. For example, we believe that with these last
inferences, \Metis summarized many essential steps in the conversion
or deduction from one formula or a list of formulas to another.
That is inconvenient since, our proof reconstruction approach needs
take into account any assumption or consideration as ordering for
literals, application of simplification theorems, reductions, among
other valid transformations.

\unsure{Should I give an Example?}

In this part, our main purpose consists of giving a response to
these rules looking at the \Metis' source code while inspecting a
representative set of solutions of \CPL problems
\cite{Prieto-Cubides2017}generated by \Metis.
This is an attempt to fulfill the \Metis's documentation,
describing our understanding of each rule, to later on in section
\ref{sec:proof-reconstruction}, formalize them effectively in \Agda.

\vskip 2mm

% ...................................................................

@@

% ...................................................................

@@

% ...................................................................

\textit{Normalization.} \Metis uses the \canonicalize rule to
normalize a propositional formula that comes from an axiom or a
definition. It often converts each expression in one of its normal
form, the conjunctive normal form, the negative normal form or the
disjunctive normal form. Posterior, this rule simplifies the formula
with some definitions, applying recursively on the structure's
formula conjunctions and disjunctions simplifications to remove
tautologies (see Fig.~\ref{fig:conjunctive-disjunctive-simpl} with a
list of these theorems. We had assumed the commutative property for
conjunction and disjunction connectives).

@@

\[%\scalebox{0.9}{
  \begin{bprooftree}
    \AxiomC{$\varphi \vee \bot$}
    \UnaryInfC{$\varphi$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{$\varphi \vee \top$}
    \UnaryInfC{$\top$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{$\varphi \vee \neg \varphi$}
    \UnaryInfC{$\top$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{$\varphi \vee \varphi$}
    \UnaryInfC{$\varphi$}
  \end{bprooftree}
%}
\]
\caption{Defintions of \canonicalize inference.}
\label{fig:conjunctive-disjunctive-simpl}
\end{figure}

% ...................................................................

@@

% ...................................................................

\textit{Simplification.}
The \simplify rule is a $n$-ary inference that performs
simplification of definitions. This rule transverses a list of
formulas by applying different theorems (e.g., \emph{modus pones},
\emph{modus tollens}, or \emph{disjunctive syllogism}) including the
list of theorems used by \canonicalize in
Fig.\ref{fig:conjunctive-disjunctive-simpl} and the resolution
theorem of \resolve rule. Moreover, many things happen inside this
\Metis' procedure, and it is somewhat complicated to grasp it
completely.

### 2017-09-07 paper/sections/related-work.tex
[ paper ] major changes, abstract, conclusions was added.

\name{Sledgehammer} is a tool for \name{Isabelle} proof assistant
~\cite{paulson1994isabelle} that provides a full integration of automatic
theorem provers including \ATPs
(see, for example, \cite{meng2006automation}, \cite{blanchette2013extending}
and \cite{Fleury2014}) and \SMT solvers (see, for instance,
\cite{hurlin07practical}, \cite{bohme2010},
\cite{blanchette2013extending}, and \cite{Fleury2014}) with
\name{Isabelle/HOL} \cite{nipkow2002isabelle}, the specialization of
\name{Isabelle} for higher-order logic. Sultana, Benzm{\"{u}}ller, and
Paulson~\cite{Een2004} integrates \name{Leo-II} and \name{Satallax}, two
theorem provers for high-order logic with \name{Isabelle/HOL} proposing a
modular proof reconstruction work flow.

@@

%(Vol. 36, pp. 127–136).

@@

% Given a conjecture, it attempts to find suitable premises, then calls
% external ATPs such as E[Sch13], Vampire [KV13], or Z3 [dMB08], and
% attempts to reconstruct the proof using the premises used by the ATP.
% To reconstruct proofs, it uses tactics such as MESON, simplification,
% and a
% few other decision procedures, however, these are sometimes not p
% owerful
% enough to reconstruct proofs found by the external ATPs.

\name{Waldmeister} is an automatic theorem prover for unit equational
logic \cite{hillenbrand1997}.
Foster and Struth~\cite{foster2011integrating} integrate \name{Waldmeister} into
\Agda~\cite{agdateam}. This integration requires a proof
reconstruction step, but authors' approach is restricted to pure
equational logic --also called identity theory~\cite{humberstone2011}--
that is, first-order logic with equality but no other predicate symbols
and no functions symbols~\cite{appel1959}.

Kanso and Setzer~\cite{kanso2016light} integrate \name{Z3} in \Agda
but they also integrate the propositional fragment of the \name{E} prover in
\cite{Kanso2012}. They cataloged these two integration as
\emph{Oracle and Reflection} and \emph{Oracle and Justification}, respectively.
Their integration with the \name{E} prover is the most related work with our
approach found at the moment while we write this document. Besides the use
of the same \ITP, crucial differences easily arise like decidability of
propositional logic using semantics that we do not take into account for the
reason we describe later on in section~\ref{sec:conclusions}.

>>>>

\name{Sledgehammer} is a tool for \name{Isabelle} proof assistant
\cite{paulson1994isabelle} that provides a full integration of
automatic theorem provers including \ATPs (see, for example,
\cite{meng2006automation}, \cite{blanchette2013extending} and
\cite{Fleury2014}) and \SMT solvers (see, for instance,
\cite{hurlin07practical}, \cite{bohme2010}, \cite{blanchette2013extending}, and \cite{Fleury2014}) with
\name{Isabelle/HOL} \cite{nipkow2002isabelle}, the specialization of
\name{Isabelle} for higher-order logic. Sultana, Benzm{\"{u}}ller,
and Paulson~\cite{Een2004} integrates \name{Leo-II} and
\name{Satallax}, two theorem provers for high-order logic with
\name{Isabelle/HOL} proposing a modular proof reconstruction work
flow.

@@

% (Vol. 36, pp. 127–136).

@@

% Given a conjecture, it attempts to find suitable premises, then
% calls external ATPs such as E[Sch13], Vampire [KV13], or Z3 [dMB08],
% and attempts to reconstruct the proof using the premises used by the
% ATP. To reconstruct proofs, it uses tactics such as MESON,
% simplification, and a few other decision procedures, however, these
% are sometimes not p owerful enough to reconstruct proofs found by
% the external ATPs.

\name{Waldmeister} is an automatic theorem prover for unit
equational logic \cite{hillenbrand1997}. Foster and Struth
\cite{foster2011integrating} integrate \name{Waldmeister} into \Agda
\cite{agdateam}. This integration requires a proof reconstruction
step, but authors' approach is restricted to pure equational logic
--also called identity theory~\cite{humberstone2011}-- that is,
first-order logic with equality but no other predicate symbols and
no functions symbols~\cite{appel1959}.

% Kanso and Setzer~\cite{kanso2016light} integrate \name{Z3} in \Agda
% but they also integrate the propositional fragment of the \name{E} prover in
% \cite{Kanso2012}. They cataloged these two integration as
% \emph{Oracle and Reflection} and \emph{Oracle and Justification}, respectively.
% Their integration with the \name{E} prover is the most related work with our
% approach found at the moment while we write this document. Besides the use
% of the same \ITP, crucial differences easily arise like decidability of
% propositional logic using semantics that we do not take into account for the
% reason we describe later on in section~\ref{sec:conclusions}.
